Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal to Beyond

### 2023-12-16 main.tex
:pencil: Add MIMO IC

% glossaries-extra
\glsdisablehyper
\setabbreviationstyle[acronym]{long-short}
\newacronym{ao}{AO}{Alternating Optimization}
\newacronym{bd}{BD}{Beyond-Diagonal}
\newacronym{dof}{DoF}{Degree of Freedom}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{rcg}{RCG}{Riemannian Conjugate Gradient}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{pc}{PC}{Point-to-point Channel}
\newacronym{ic}{IC}{Interference Channel}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}

@@

	\begin{subsection}{Rate Maximization}
		The problem is formulated w.r.t. precoder (instead of transmit covariance matrix) for reference:
		\begin{maxi!}
			{\scriptstyle{\mathbf{W},\mathbf{\Theta}}}{\log \det \biggl(\mathbf{I} + \frac{\mathbf{W}^\mathsf{H}\mathbf{H}^\mathsf{H}\mathbf{H}\mathbf{W}}{\sigma_n^2}\biggr)}{\label{op:pc_rate}}{\label{ob:pc_rate}}
			\addConstraint{\lVert \mathbf{W} \rVert _\mathrm{F}^2 \le P}{}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g.}{}{\label{cs:pc_rate_unitary}}
		\end{maxi!}

@@

		We then analyze the channel shaping \emph{capability} of \gls{bd} \gls{ris} under specific setups.
		\begin{subsubsection}{Rank-Deficient Channel}
			In rank-deficient channels, \gls{bd} \gls{ris} $\mathbf{\Theta}^\mathrm{B}$ cannot achieve a higher \gls{dof} than diagonal \gls{ris} $\mathbf{\Theta}^\mathrm{D}$.
			This is because $\mathrm{sv}(\mathbf{\Theta}^\mathrm{B}) = \mathrm{sv}(\mathbf{\Theta}^\mathrm{D}) = \boldsymbol{1}$ and
			\begin{equation}
				\begin{split}
					\mathrm{rank}(\mathbf{H})
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \mathrm{rank}(\mathbf{H}^\mathrm{B} \mathbf{\Theta} \mathbf{H}^\mathrm{F}) \\
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \min \bigl( \mathrm{rank}(\mathbf{H}^\mathrm{B}), \mathrm{rank}(\mathbf{\Theta}), \mathrm{rank}(\mathbf{H}^\mathrm{F}) \bigr).
				\end{split}
			\end{equation}
			% However, \gls{bd} \gls{ris} can still guarantee a higher rate at finite \gls{snr}.
			Note \gls{bd} \gls{ris} can still provide a higher indirect \gls{snr} as shown in Fig.~\ref{sm:pc_power_sx} and \ref{sm:pc_power_bond}.
		\end{subsubsection}

@@

		\begin{subsubsection}{Fully-Connected \glsentryshort{ris} Without Direct Link}
			Denote the singular value decomposition of forward / backward channels as $\mathbf{H}^{\mathrm{B}/\mathrm{F}} = \mathbf{U}^{\mathrm{B}/\mathrm{F}} \mathbf{\Sigma}^{\mathrm{B}/\mathrm{F}} {\mathbf{V}^{\mathrm{B}/\mathrm{F}}}^\mathsf{H}$.
			The composite channel is
			\begin{equation}
				\mathbf{H} = \mathbf{H}^\mathrm{B} \mathbf{\Theta} \mathbf{H}^\mathrm{F} = \mathbf{U}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{X} \mathbf{\Sigma}^\mathrm{F} {\mathbf{V}^\mathrm{F}}^\mathsf{H},
			\end{equation}
			where $\mathbf{X} = {\mathbf{V}^\mathrm{B}}^\mathsf{H} \mathbf{\Theta} \mathbf{U}^\mathrm{F}$.
			% Since unitary matrices are closed under multiplication,
			\begin{proposition}
				In this case, the singular value bounds on $\mathbf{H}$ are equivalent to the singular value bounds on $\mathbf{BF}$, where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with singular values $\mathbf{\Sigma}^\mathrm{B}$ and $\mathbf{\Sigma}^\mathrm{F}$.
				% The only singular value bounds applied here are those apply to the singular value bound of $\mathbf{F}\mathbf{B}$
			\end{proposition}
			\begin{proof}
				We first observe that singular value control problem can be solved w.r.t. unitary $\mathbf{X}$ and retrieved by $\mathbf{\Theta} = \mathbf{V}^\mathrm{B} \mathbf{X} {\mathbf{U}^\mathrm{F}}^\mathsf{H}$.
				Also, $\mathrm{sv}(\mathbf{U}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{X} \mathbf{\Sigma}^\mathrm{F} {\mathbf{V}^\mathrm{F}}^\mathsf{H}) = \mathrm{sv}(\bar{\mathbf{U}}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{\bar{V}}{{}^\mathrm{B}}^\mathsf{H} \bar{\mathbf{U}}^\mathrm{F} \mathbf{\Sigma}^\mathrm{F} \mathbf{\bar{V}}{{}^\mathrm{F}}^\mathsf{H}) = \mathrm{sv}(\mathbf{BF})$ where $\bar{\mathbf{U}}^{\mathrm{B}/\mathrm{F}}$ and $\bar{\mathbf{V}}^{\mathrm{B}/\mathrm{F}}$ are arbitrary unitary matrices.
			\end{proof}
			The problem now becomes, given $\mathbf{\Sigma}^\mathrm{B}$ and $\mathbf{\Sigma}^\mathrm{F}$, what can we say about the singular value of $\mathbf{BF}$.
			One comprehensive answer is Horn's inequality \cite{Bhatia2001}: for all admissible triples $(I, J, K)$,
			\begin{equation}
				\prod_{k \in {K}} \sigma_k(\mathbf{BF}) \leq \prod_{i \in {I}} \sigma_i(\mathbf{B}) \prod_{j \in {J}} \sigma_j(\mathbf{F}).
			\end{equation}
			It gives upper bound on the largest singular value and lower bound on the smallest singular value:
			\begin{align}
				\sigma_1(\mathbf{BF}) \le \sigma_1(\mathbf{B}) \sigma_1(\mathbf{F}) \label{iq:singular_value_largest}\\
				\sigma_N(\mathbf{BF}) \geq \sigma_N(\mathbf{B}) \sigma_N(\mathbf{F}) \label{iq:singular_value_smallest}.
			\end{align}
			Another useful result is introduced in \cite{Hogben2013}: for all $p>0$,
			\begin{equation}
				\sum_n \sigma_n^p(\mathbf{BF}) \leq \sum_n \sigma_n^p(\mathbf{B}) \sigma_n^p(\mathbf{F}) \label{iq:singular_value_power_sum}.
			\end{equation}
			When $p=2$, it implies the channel energy is upper bounded by the sum of element-wise power product of the forward and backward channels, as illustrated in Fig.~\subref*{sm:pc_power_bond_nd}.
			Interestingly, \eqref{iq:singular_value_largest}--\eqref{iq:singular_value_power_sum} are simultaneously tight when $\mathbf{X} = \mathbf{I}$ and $\mathbf{\Theta} = \mathbf{V}^\mathrm{B} {\mathbf{U}^\mathrm{F}}^\mathsf{H}$.
			This solution was claimed in \cite{Bartoli2023} to achieve channel capacity, but it is not true at moderate \gls{snr}.
			% In \cite{Bartoli2023}, the authors claimed this solution also achieves channel capacity but it is
		\end{subsubsection}

		Finally, we characterize the \emph{Pareto front} of channel singular values via optimization approach.
	\end{subsection}
\end{section}

>>>>

% glossaries-extra
\glsdisablehyper
\setabbreviationstyle[acronym]{long-short}
\newacronym{ao}{AO}{Alternating Optimization}
\newacronym{bd}{BD}{Beyond-Diagonal}
\newacronym{bcd}{BCD}{Block Coordinate Descent}
\newacronym{dof}{DoF}{Degree of Freedom}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{rcg}{RCG}{Riemannian Conjugate Gradient}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{pc}{PC}{Point-to-point Channel}
\newacronym{ic}{IC}{Interference Channel}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}
\newacronym{wsr}{WSR}{Weighted Sum-Rate}
\newacronym{mmse}{MMSE}{Minimum Mean-Square Error}
\newacronym{wmmse}{WMMSE}{Weighted \gls{mmse}}
\newacronym{mse}{MSE}{Mean-Square Error}

@@

	\begin{subsection}{Rate Maximization}
		The problem is formulated w.r.t. precoder (instead of transmit covariance matrix) for reference:
		\begin{maxi!}
			{\scriptstyle{\mathbf{W},\mathbf{\Theta}}}{R = \log \det \biggl(\mathbf{I} + \frac{\mathbf{W}^\mathsf{H}\mathbf{H}^\mathsf{H}\mathbf{H}\mathbf{W}}{\eta}\biggr)}{\label{op:pc_rate}}{\label{ob:pc_rate}}
			\addConstraint{\lVert \mathbf{W} \rVert _\mathrm{F}^2 \le P}{}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g.}{}{\label{cs:pc_rate_unitary}}
		\end{maxi!}

@@

		We then analyze the channel shaping \emph{capability} of \gls{bd} \gls{ris} under specific setups.
		\begin{subsubsection}{Rank-Deficient Channel}
			\label{sc:pc_rank_deficient}
			In rank-deficient channels, \gls{bd} \gls{ris} $\mathbf{\Theta}^\mathrm{B}$ cannot achieve a higher \gls{dof} than diagonal \gls{ris} $\mathbf{\Theta}^\mathrm{D}$.
			This is because $\mathrm{sv}(\mathbf{\Theta}^\mathrm{B}) = \mathrm{sv}(\mathbf{\Theta}^\mathrm{D}) = \boldsymbol{1}$ and
			\begin{equation}
				\begin{split}
					\mathrm{rank}(\mathbf{H})
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \mathrm{rank}(\mathbf{H}^\mathrm{B} \mathbf{\Theta} \mathbf{H}^\mathrm{F}) \\
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \min \bigl( \mathrm{rank}(\mathbf{H}^\mathrm{B}), \mathrm{rank}(\mathbf{\Theta}), \mathrm{rank}(\mathbf{H}^\mathrm{F}) \bigr).
				\end{split}
			\end{equation}
			% However, \gls{bd} \gls{ris} can still guarantee a higher rate at finite \gls{snr}.
			Note \gls{bd} \gls{ris} can still provide a higher indirect \gls{snr} as shown in Fig.~\ref{sm:pc_power_sx} and \ref{sm:pc_power_bond}.
		\end{subsubsection}

@@

		\begin{subsubsection}{Fully-Connected \glsentryshort{ris} Without Direct Link}
			Denote the singular value decomposition of forward / backward channels as $\mathbf{H}^{\mathrm{B}/\mathrm{F}} = \mathbf{U}^{\mathrm{B}/\mathrm{F}} \mathbf{\Sigma}^{\mathrm{B}/\mathrm{F}} {\mathbf{V}^{\mathrm{B}/\mathrm{F}}}^\mathsf{H}$.
			The composite channel is
			\begin{equation}
				\mathbf{H} = \mathbf{H}^\mathrm{B} \mathbf{\Theta} \mathbf{H}^\mathrm{F} = \mathbf{U}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{X} \mathbf{\Sigma}^\mathrm{F} {\mathbf{V}^\mathrm{F}}^\mathsf{H},
			\end{equation}
			where $\mathbf{X} = {\mathbf{V}^\mathrm{B}}^\mathsf{H} \mathbf{\Theta} \mathbf{U}^\mathrm{F}$.
			% Since unitary matrices are closed under multiplication,
			\begin{proposition}
				In this case, the singular value bounds on $\mathbf{H}$ are equivalent to the singular value bounds on $\mathbf{BF}$, where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with singular values $\mathbf{\Sigma}^\mathrm{B}$ and $\mathbf{\Sigma}^\mathrm{F}$.
				% The only singular value bounds applied here are those apply to the singular value bound of $\mathbf{F}\mathbf{B}$
			\end{proposition}
			\begin{proof}
				We first observe that singular value control problem can be solved w.r.t. unitary $\mathbf{X}$ and retrieved by $\mathbf{\Theta} = \mathbf{V}^\mathrm{B} \mathbf{X} {\mathbf{U}^\mathrm{F}}^\mathsf{H}$.
				Also, $\mathrm{sv}(\mathbf{U}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{X} \mathbf{\Sigma}^\mathrm{F} {\mathbf{V}^\mathrm{F}}^\mathsf{H}) = \mathrm{sv}(\bar{\mathbf{U}}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{\bar{V}}{{}^\mathrm{B}}^\mathsf{H} \bar{\mathbf{U}}^\mathrm{F} \mathbf{\Sigma}^\mathrm{F} \mathbf{\bar{V}}{{}^\mathrm{F}}^\mathsf{H}) = \mathrm{sv}(\mathbf{BF})$ where $\bar{\mathbf{U}}^{\mathrm{B}/\mathrm{F}}$ and $\bar{\mathbf{V}}^{\mathrm{B}/\mathrm{F}}$ are arbitrary unitary matrices.
			\end{proof}
			The problem now becomes, given $\mathbf{\Sigma}^\mathrm{B}$ and $\mathbf{\Sigma}^\mathrm{F}$, what can we say about the singular value of $\mathbf{BF}$.
			One comprehensive answer is Horn's inequality \cite{Bhatia2001}: for all admissible triples $(I, J, K)$,
			\begin{equation}
				\prod_{k \in {K}} \sigma_k(\mathbf{BF}) \leq \prod_{i \in {I}} \sigma_i(\mathbf{B}) \prod_{j \in {J}} \sigma_j(\mathbf{F}).
			\end{equation}
			It gives upper bound on the largest singular value and lower bound on the smallest singular value:
			\begin{align}
				\sigma_1(\mathbf{BF}) \le \sigma_1(\mathbf{B}) \sigma_1(\mathbf{F}) \label{iq:singular_value_largest}\\
				\sigma_N(\mathbf{BF}) \geq \sigma_N(\mathbf{B}) \sigma_N(\mathbf{F}) \label{iq:singular_value_smallest}.
			\end{align}
			Another useful result is introduced in \cite{Hogben2013}: for all $p>0$,
			\begin{equation}
				\sum_n \sigma_n^p(\mathbf{BF}) \leq \sum_n \sigma_n^p(\mathbf{B}) \sigma_n^p(\mathbf{F}) \label{iq:singular_value_power_sum}.
			\end{equation}
			When $p=2$, it implies the channel energy is upper bounded by the sum of element-wise power product of the forward and backward channels, as illustrated in Fig.~\subref*{sm:pc_power_bond_nd}.
			Interestingly, \eqref{iq:singular_value_largest}--\eqref{iq:singular_value_power_sum} are simultaneously tight when $\mathbf{X} = \mathbf{I}$ and $\mathbf{\Theta} = \mathbf{V}^\mathrm{B} {\mathbf{U}^\mathrm{F}}^\mathsf{H}$.
			This solution was claimed in \cite{Bartoli2023} to achieve channel capacity, but it is not true at moderate \gls{snr}.
		\end{subsubsection}

		Finally, we characterize the \emph{Pareto front} of channel singular values via optimization approach.
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{J_1 = \sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pc_singular}}{\label{ob:pc_singular}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g,}{}{}
		\end{maxi!}
		where $\rho_n$ is the weight of $n$-th singular value.
		The complex derivative of \eqref{ob:pc_singular} w.r.t. \gls{ris} block $g$ is
		\begin{equation}
			\frac{\partial J_1}{\partial \mathbf{\Theta}_g^*} = {\mathbf{H}_g^\mathrm{B}}^\mathsf{H} \mathbf{U} \mathrm{diag}(\boldsymbol{\rho}) \mathbf{V}^\mathsf{H} {\mathbf{H}_g^\mathrm{F}}^\mathsf{H},
			\label{eq:pc_singular_gradient_ris}
		\end{equation}
		where $\mathbf{U}$ and $\mathbf{V}$ are left and right singular matrix of $\mathbf{H}$.
		\eqref{op:pc_singular} can be solved by \gls{rcg} Algorithm~\ref{ag:pc_rate_ris} with \eqref{eq:pc_rate_gradient_ris} replaced by \eqref{eq:pc_singular_gradient_ris}.

		\begin{figure}[!t]
			\centering
			\subfloat[Pareto Front\label{sm:pc_singular_pareto_bond}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_pareto_bond.tex}
				}
			}
			\subfloat[Evolving Trend\label{sm:pc_singular_trend_bond}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_singular_trend_bond.tex}
				}
			}
			\caption{Singular value Pareto front and trend for $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (2, 16, 2)$.}
			\label{sm:pc_singular_bond}
		\end{figure}
		The Pareto front and evolving trend of channel singular values are shown in Fig.~\subref*{sm:pc_singular_pareto_bond} and \subref*{sm:pc_singular_trend_bond}.
		Clearly, \gls{bd} \gls{ris} with a larger group size can redistribute the channel singular values to a wider range.
	\end{subsection}
\end{section}

\begin{section}{\glsentryshort{mimo}-\glsentryshort{ic}}
	\begin{subsection}{Leakage Interference Minimization}
		\begin{mini!}
			{\scriptstyle{\mathbf{\Theta}, \{\mathbf{G}_k\}, \{\mathbf{W}_k\}}}{\mathop{\sum\sum}_{j \neq k} \left\lVert \mathbf{G}_k (\mathbf{H}_{kj}^\mathrm{D} + \mathbf{H}_k^\mathrm{B} \mathbf{\Theta} \mathbf{H}_j^\mathrm{F}) \mathbf{W}_j \right\rVert _{\mathrm{F}}^2}{}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g}{}{}
			\addConstraint{\mathbf{G}_k \mathbf{G}_k^\mathsf{H}=\mathbf{I}, \quad \mathbf{W}_k^\mathsf{H} \mathbf{W}_k=\mathbf{I}, \quad \forall k.}{}{}
		\end{mini!}
		The non-convex problem can be solved by \gls{bcd} method.
		For a given $\mathbf{\Theta}$, it reduces to conventional linear beamforming problem, for which an iterative algorithm alternating between the original and reciprocal networks is proposed in \cite{Gomadam2011,Clerckx2013}.
		At iteration $r$, the combiner at receiver $k$ is updated as
		\begin{equation}
			\mathbf{G}_k^{(r)} = {\mathbf{U}_{k,N}^{(r-1)}}^\mathsf{H},
		\end{equation}
		where $\mathbf{U}_{k,N}^{(r-1)}$ is the eigenvectors corresponding to $N$ smallest eigenvalues of interference covariance matrix $\mathbf{Q}_k^{(r-1)} = \sum_{j \ne k} \mathbf{H}_{kj} \mathbf{W}_j^{(r-1)} {\mathbf{W}_j^{(r-1)}}^\mathsf{H} \mathbf{H}_{kj}^\mathsf{H}$.
		The precoder at transmitter $j$ is updated as
		\begin{equation}
			\mathbf{W}_j^{(r)} = \bar{\mathbf{U}}_{j,N}^{(r)},
		\end{equation}
		where $\bar{\mathbf{U}}_{j,N}^{(r)}$ corresponds to interference covariance matrix $\bar{\mathbf{Q}}_j^{(r)} = \sum_{k \ne j} \mathbf{H}_{kj}^\mathsf{H} {\mathbf{G}_k^{(r)}}^\mathsf{H} \mathbf{G}_k^{(r)} \mathbf{H}_{kj}$ in the reciprocal network.
		Once $\{\mathbf{G}_k\}$ and $\{\mathbf{W}_k\}$ are determined, we define $\bar{\mathbf{H}}_{kj}^\mathrm{D} \triangleq \mathbf{G}_k \mathbf{H}_{kj}^\mathrm{D} \mathbf{W}_j$, $\bar{\mathbf{H}}_k^\mathrm{B} \triangleq \mathbf{G}_k \mathbf{H}_k^\mathrm{B}$, and $\bar{\mathbf{H}}_j^\mathrm{F} \triangleq \mathbf{H}_j^\mathrm{F} \mathbf{W}_j$.
		The \gls{bd} \gls{ris} subproblem reduces to
		\begin{mini!}
			{\scriptstyle{\mathbf{\Theta}}}{\mathop{\sum\sum}_{j \neq k} \left\lVert (\bar{\mathbf{H}}_{kj}^\mathrm{D} + \bar{\mathbf{H}}_k^\mathrm{B} \mathbf{\Theta} \bar{\mathbf{H}}_j^\mathrm{F}) \right\rVert _{\mathrm{F}}^2}{\label{op:ic_interference_ris}}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g.}{}{}
		\end{mini!}

		\begin{proposition}
			Start from any $\mathbf{\Theta}^{(0)}$, the sequence
			\begin{equation}
				\mathbf{\Theta}_g^{(r+1)} = \mathbf{U}_g^{(r)} \mathbf{V}_g^{(r)}, \quad \forall g
			\end{equation}
			converges to a stationary point of \eqref{op:ic_interference_ris}, where $\mathbf{U}_g^{(r)}$ and $\mathbf{V}_g^{(r)}$ are left and right singular matrix of
			\begin{equation}
				\mathbf{M}_g^{(r)} = \mathop{\sum\sum}_{j \neq k} \bigl(\mathbf{B}_{k,g} \mathbf{\Theta}_g^{(r)} \mathbf{H}_{j,g}^\mathrm{F} - {\mathbf{H}_{k,g}^\mathrm{B}}^\mathsf{H} \mathbf{D}_{kj,g}^{(r)}\bigr) {\mathbf{H}_{j,g}^\mathrm{F}}^\mathsf{H},
			\end{equation}
			where $\mathbf{B}_{k,g} = \lambda_1\bigl(\mathbf{H}_{k,g}^\mathrm{B} {\mathbf{H}_{k,g}^\mathrm{B}}^\mathsf{H}\bigr) \mathbf{I} - \mathbf{H}_{k,g}^\mathrm{B} {\mathbf{H}_{k,g}^\mathrm{B}}^\mathsf{H}$ and
			\begin{equation}
				\mathbf{D}_{kj,g}^{(r)} = \mathbf{H}_{jk}^\mathrm{D} + \sum_{g'<g} {\mathbf{H}_{k,g'}^\mathrm{B}}^\mathsf{H} \mathbf{\Theta}_{g'}^{(r+1)} \mathbf{H}_{k,g'}^\mathrm{F} + \sum_{g'>g} {\mathbf{H}_{k,g'}^\mathrm{B}}^\mathsf{H} \mathbf{\Theta}_{g'}^{(r)} \mathbf{H}_{k,g'}^\mathrm{F}.
			\end{equation}
		\end{proposition}
		\begin{proof}
			To be added.
		\end{proof}

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/ic_interference_sx.tex}
			}
			\caption{Average leakage interference versus \gls{ris} elements $N^\mathrm{S}$ and group size $L$ for $(N^\mathrm{T}, N^\mathrm{R}, N^\mathrm{E}, K) = (8, 4, 3, 5)$. Transmitters and receivers are randomly generated in a disk of radius 50 m centered at the \gls{ris}. Reference pathloss at \qty{1}{\meter} is \qty{-30}{\dB} and $(\gamma^\mathrm{D}, \gamma^\mathrm{F}, \gamma^\mathrm{B}) = (3, 2.4, 2.4)$.}
			\label{sm:ic_interference_sx}
		\end{figure}
		Fig.~\ref{sm:ic_interference_sx} illustrates how \gls{bd} \gls{ris} helps to reduce the leakage interference.
		In this case, a fully-connected $2^n$-element \gls{bd} \gls{ris} is almost as good as a diagonal $2^{n+2}$-element \gls{ris} in terms of leakage interference.
		Interestingly, the result suggests that \gls{bd} \gls{ris} can achieve a higher \gls{dof} than diagonal \gls{ris} in \gls{mimo}-\gls{ic}, which is not the case in \gls{mimo}-\gls{pc} (as discussed in \ref{sc:pc_rank_deficient}).
	\end{subsection}

	\begin{subsection}{Weighted Sum-Rate Maximization}
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}, \{\mathbf{W}_k\}}}{J_2 = \sum_k \rho_k \log \det \biggl(\mathbf{I} + \mathbf{W}_k \mathbf{H}_{kj}^\mathsf{H} \mathbf{Q}_k^{-1} \mathbf{H}_{kj} \mathbf{W}_k\biggr)}{\label{op:ic_rate}}{\label{ob:ic_rate}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g}{}{}
			\addConstraint{\lVert \mathbf{W}_k \rVert _\mathrm{F}^2 \le P_k. \quad \forall k}{}{}
		\end{maxi!}
		where $\rho_k$ is the weight of user $k$ and $\mathbf{Q}_k$ is the interference-plus-noise covariance matrix
		\begin{equation}
			\mathbf{Q}_k = \sum_{j \ne k} \mathbf{H}_{kj} \mathbf{W}_j {\mathbf{W}_j}^\mathsf{H} \mathbf{H}_{kj}^\mathsf{H} + \eta \mathbf{I}.
		\end{equation}
		For a given $\mathbf{\Theta}$, \eqref{op:ic_rate} reduces to conventional linear beamforming problem, for which a closed-form iterative solution based on \gls{wsr}-\gls{wmmse} relationship is proposed in \cite{Negro2010}.
		At iteration $r$, the \gls{mmse} combiner at receiver $k$ is
		\begin{equation}
			\mathbf{G}_k^{(r)} = {\mathbf{W}_k^{(r-1)}}^\mathsf{H} \mathbf{H}_{kk}^\mathsf{H} \bigl(\mathbf{Q}_k^{(r-1)} + \mathbf{H}_{kk} \mathbf{W}_k^{(r-1)} {\mathbf{W}_k^{(r-1)}}^\mathsf{H} \mathbf{H}_{kk}^\mathsf{H}\bigr)^{-1},
		\end{equation}
		the corresponding error matrix is
		\begin{equation}
			\mathbf{E}_k^{(r)} = \bigl(\mathbf{I} + {\mathbf{W}_k^{(r-1)}}^\mathsf{H} \mathbf{H}_{kk}^\mathsf{H} \mathbf{Q}_k^{(r-1)} \mathbf{H}_{kk} \mathbf{W}_k^{(r-1)}\bigr)^{-1},
		\end{equation}
		the \gls{mse} weight is
		\begin{equation}
			\mathbf{\Omega}_k^{(r)} = \rho_k {\mathbf{E}_k^{(r)}}^{-1},
		\end{equation}
		the Lagrange multiplier is
		% \begin{equation}
		% 	\begin{split}
		% 		\lambda_k^{(r)}
		% 		& = \frac{\sum_{j \ne k} \mathrm{tr}(\mathbf{\Omega}_k^{(r)}\mathbf{F}_{kj}^{(r)} {\mathbf{F}_{kj}^{(r)}}^\mathsf{H}) - \sum_{j \ne k} \mathrm{tr}(\mathbf{\Omega}_j^{(r)}\mathbf{F}_{jk}^{(r)} {\mathbf{F}_{jk}^{(r)}}^\mathsf{H})}{P_k} \\
		% 		& \quad + \frac{\mathrm{tr}(\eta \mathbf{\Omega}_k^{(r)} \mathbf{G}_k^{(r)}{\mathbf{G}_k^{(r)}}^\mathsf{H})}{P_k}
		% 	\end{split}
		% \end{equation}
		\begin{equation}
			\lambda_k^{(r)} = \frac{\mathrm{tr}\bigl(\eta \mathbf{\Omega}_k^{(r)} \mathbf{G}_k^{(r)}{\mathbf{G}_k^{(r)}}^\mathsf{H} + \sum_j \mathbf{\Omega}_k^{(r)}\mathbf{T}_{kj}^{(r)} {\mathbf{T}_{kj}^{(r)}}^\mathsf{H} - \mathbf{\Omega}_j^{(r)}\mathbf{T}_{jk}^{(r)} {\mathbf{T}_{jk}^{(r)}}^\mathsf{H} \bigr)}{P_k},
		\end{equation}
		where $\mathbf{T}_{kj}^{(r)} = \mathbf{G}_k^{(r)} \mathbf{H}_{kj} \mathbf{W}_j^{(r)}$.
		The precoder at transmitter $k$ is
		\begin{equation}
			\mathbf{W}_k^{(r)} = \Bigl(\sum_j \mathbf{H}_{jk}^\mathsf{H} {\mathbf{G}_j^{(r)}}^\mathsf{H} \mathbf{\Omega}_k^{(r)} \mathbf{G}_j^{(r)} \mathbf{H}_{jk} + \lambda_k^{(r)} \mathbf{I} \Bigr)^{-1} \mathbf{H}_{kk}^\mathsf{H} {\mathbf{G}_j^{(r)}}^\mathsf{H} \mathbf{\Omega}_k^{(r)}.
		\end{equation}
		Once $\{\mathbf{W}_k\}$ is determined, the complex derivative of \eqref{ob:ic_rate} w.r.t. \gls{ris} block $g$ is
		\begin{equation}
			\begin{split}
				\frac{\partial J_2}{\partial \mathbf{\Theta}_g^*}
				& = \sum_k \rho_k {\mathbf{H}_{k,g}^\mathrm{B}}^\mathsf{H} \mathbf{Q}_k^{-1} \mathbf{H}_{kk} \mathbf{W}_k \mathbf{E}_k \mathbf{W}_k^\mathsf{H} \\
				& \quad \times \bigl({\mathbf{H}_{k,g}^\mathrm{F}}^\mathsf{H} - \mathbf{H}_{kk}^\mathsf{H} \mathbf{Q}_k^{-1} \sum_{j \ne k} \mathbf{H}_{kj} \mathbf{W}_j \mathbf{W}_j^\mathsf{H} {\mathbf{H}_{j,g}^\mathrm{F}}^\mathsf{H}\bigr).
			\end{split}
			\label{eq:ic_rate_gradient_ris}
		\end{equation}
		The \gls{ris} subproblem can be solved by \gls{rcg} Algorithm~\ref{ag:pc_rate_ris} with \eqref{eq:pc_rate_gradient_ris} replaced by \eqref{eq:ic_rate_gradient_ris}.

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/ic_rate_sx.tex}
			}
			\caption{Average weighted sum-rate versus \gls{ris} elements $N^\mathrm{S}$ and group size $L$ for $(N^\mathrm{T}, N^\mathrm{R}, N^\mathrm{E}, K) = (8, 4, 3, 5)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$, $\rho_k = 1, \ \forall k$.}
			\label{sm:ic_rate_sx}
		\end{figure}
		A new observation from Fig.~\ref{sm:ic_rate_sx} that the interference alignment capability of \gls{bd} \gls{ris} scales much faster with group size than number of elements.\footnote{The results are not very stable and depend heavily on initialization.}
	\end{subsection}
\end{section}
